---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
data <- read.csv("activity.csv",stringsAsFactors=FALSE)
data$date <- as.Date(data$date)
summary(data)
```


## What is mean total number of steps taken per day?

```{r}
days <- group_by(data, date)
days <- summarise(days, nintervals=n(), avgsteps=mean(steps,na.rm=TRUE), totsteps=nintervals*avgsteps)
plot(days$date, days$totsteps, type="h",xlab="Date",ylab="Total Steps per Day",main="Histogram - Total Number of Steps per Day")
```
 
 __Total Number of Steps per Day__  
 __Mean:__ `r sprintf("%.2f", mean(days$totsteps, na.rm=TRUE))`  
 __Median:__ `r sprintf("%.0f", median(days$totsteps, na.rm=TRUE))`  


## What is the average daily activity pattern?

```{r}
days <- group_by(data, interval)
days <- summarise(days, avgsteps=mean(steps,na.rm=TRUE))
days <- mutate(days, minutes=(interval/100), seconds=(interval - 100*minutes + 60*minutes), index=seconds/5)
plot(days$index, days$avgsteps, type="l", xlab="Interval", ylab="Average Steps", main="Daily Activity - Average Number of Steps")
```


## Imputing missing values

__Total Number of NA's:__ `r length(subset(data, is.na(steps))$steps)`

## Are there differences in activity patterns between weekdays and weekends?
